---
title: "ST558 Project2"
author: "Shan Luo, Chengxi Zhou"
date: '2022-07-03'
params:
  Channels: 'data_channel_is_bus'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
library(ggplot2)
library(tibble)
```

```{r render, echo=TRUE, eval=FALSE}
rmarkdown::render("ST558_Project2_Group10.Rmd", output_file = "lifestyle.html", params = list(Channels = 'data_channel_is_bus'))
```

# Introduction

```{r read in and subset data, echo=TRUE, eval=TRUE}
# Read in data and subset data
News <- read_csv("OnlineNewsPopularity.csv", show_col_types = FALSE) 
News <- News %>% 
  filter(!!rlang::sym(params$Channels) == 1) %>%
  select(shares, n_tokens_content, num_imgs, num_videos,
         global_rate_positive_words, global_subjectivity, is_weekend)
News$is_weekend <- factor(News$is_weekend)
```

```{r split train and test data, echo=TRUE, eval=TRUE}
# Split train and test data
set.seed(1)
trainIndex <- createDataPartition(News$shares, p = 0.7, list = FALSE)
train <- News[trainIndex, ]
test <- News[-trainIndex, ]
```

```{r check correlation, echo=TRUE, eval=TRUE}
# Check correlation of all interested variables
cor(select(News ,shares, n_tokens_content, num_imgs, num_videos,
           global_rate_positive_words, global_subjectivity))
```

If two variables have high correlation, we may think about removing one of them.  

# EDA  

## Numeric Summary Table  

```{r numeric summaries, echo=TRUE, eval=TRUE}
# Compute the summary statistics
apply(X = select(train, shares:global_subjectivity), MARGIN = 2,
      FUN = function(x) {
        summaries <- c(min(x), mean(x), median(x), max(x), sd(x), IQR(x))
        names(summaries) <- c("Min", "Mean", "Median", "Max", "Sd", "IQR")
        summaries
      })
```

From numeric summary table, if one variable's mean is greater than median, it has a right skewed distribution. If the mean is less than median, it has a left skewed distribution. If mean is close to median, it may have a symmetric distribution. If the standard deviation is unusual, there may be some outliers.  

## Contingency Table  

```{r contingency table, echo=TRUE, eval=TRUE}
# Create contingency table of predictor "is_weekend"
table(train$is_weekend)
```

From the contingency table, we can see how many articles are published on weekday and weekend.  

## Bar Plot  

```{r barplot, echo=TRUE, eval=TRUE}
# Create bar plot of predictor "is_weekend"
g <- ggplot(data = train, aes(x = is_weekend))
g + geom_bar(fill = "Red", color = "Blue") +
  labs(title = "Bar Plot of is_weekend")
```

From the bar plot, we can see how many articles are published on weekday and weekend and visualize the difference.  

```{r histograms, echo=TRUE, eval=TRUE}
# Create histogram of response "shares" and fill with predictor "is_weekend"
g <- ggplot(data = train, aes(x = shares))
g + geom_histogram(bins = 30, aes(fill = is_weekend)) +
  labs(x = "Number of Shares",
       title = "Histogram of Shares") +
  scale_fill_discrete(name = "Weekend Published", labels = c("No", "Yes"))
```

For histogram, we can see the distribution of the number of shares. If we have majority of count on the left side and less count on right side, it may have a right skewed distribution. It indicates that most of articles have small number of shares. If we have majority of count on the right side and less count on left side, it may have a left skewed distribution. It indicates that most of articles have large number of shares. If we see a bell shape, it may have a symmetric distribution. It indicating most of articles have relatively large shares.  

```{r scatterplot, echo=TRUE, eval=TRUE}
# Create scatter plot of response "shares" and predictor "n_tokens_content".
# Filled with predictor "is_weekend"
g <- ggplot(data = train, aes(x = n_tokens_content, y = shares))
g + geom_point(aes(color = is_weekend)) +
  geom_smooth(method = "lm") +
  labs(x = "Number of Words in Content",
       y = "Number of Shares",
       title = "Scatter Plot of Shares vs Number of Words in Content") +
  scale_color_discrete(name = "Weekend Published", labels = c("No", "Yes"))
```

We can inspect the trend of shares as a function of the number of words in content. If the points show an upward trend, then articles with more number of words in the content to be shared more often. If we see a negative trend then articles with more number of words in the content tend to be shared less often.

# Modeling  

## Linear Regression  

Regression models allow easy prediction of response and inference. Linear regression is that we model a response as a linear function of some predictors. Model fit by minimizing the sum of squared residuals.  

```{r first linear regression, echo=TRUE, eval=TRUE}
# Fit linear model
mod <- lm(shares ~ (n_tokens_content + num_imgs + num_videos + 
                      global_rate_positive_words + global_subjectivity + 
                      is_weekend)^2 ,data =  train)
# Use forward selection to  choose model
forward_mod <- step(mod, direction = "forward")

lmfit1 <- train(shares ~ (n_tokens_content + num_imgs + num_videos + 
                         global_rate_positive_words + global_subjectivity + 
                         is_weekend)^2,
             data = train,
             method = "lm",
             preProcess = c("center", "scale"),
             trControl = trainControl(method = "cv", number = 5))
lmfit1
lmpred1 <- predict(lmfit1, newdata = test) 
lm1 <- postResample(lmpred1, test$shares)
lm1
```

```{r second linear regression, echo=TRUE, eval=TRUE}
lmfit2 <- train(shares ~ n_tokens_content + num_imgs + num_videos + 
                         global_rate_positive_words + global_subjectivity + 
                         is_weekend,
             data = train,
             method = "lm",
             preProcess = c("center", "scale"),
             trControl = trainControl(method = "cv", number = 5))
lmfit2
lmpred2 <- predict(lmfit2, newdata = test) 
lm2 <- postResample(lmpred2, test$shares)
lm2
```

## Random Forest  

Since the response is continuous, we choose to use regression tree.  

The bootstrapping is that we resample from data or a fitted model and apply method or estimation to each resample. We see how the model or method behave.  

For Bootstrap Aggregation(Bagged) for regression tree, we create a bootstrap sample, train tree on the sample, repeat B = 1000 times, and average over these predictions as final prediction.  

Random forest follows Bootstrap Aggregation idea. We will create multiple trees from bootstrap samples and average the results. But, we will use a random subset of predictors for each bootstrap tree fit instead of using all predictors. It may make bagged trees predictions more correlated, which can help with reduction of variation.  

```{r random forest, echo=TRUE, eval=TRUE}
rffit <- train(shares ~ n_tokens_content + num_imgs + num_videos + 
                         global_rate_positive_words + global_subjectivity + 
                         is_weekend, 
               data = train, 
               method = "rf", 
               trControl = trainControl(method = "cv", number = 5), 
               preProcess = c("center", "scale"),
               tuneGrid = data.frame(mtry = 1:6))
rffit
rfpred <- predict(rffit, newdata = test) 
rf <- postResample(rfpred, test$shares)
rf
```

## Boosted Tree  

Boosting is a way that slowly trains the tree so that the tree do not over fit. For boosting, trees grow sequentially and each subsequent tree is grown on a modified version of original data. Prediction updates as trees grown.  

The process of boosted tree:  
1. Initialized prediction as 0  
2. Find residuals(observed-predicted) 
3. Fit a tree with d splits(d + 1 terminal nodes) treating the residuals as response  
4. Update predictions  
5. Update residuals for new predictions and repeat B times  

```{r boosted tree, echo=TRUE, eval=TRUE}
boostedTfit <- train(shares ~ n_tokens_content + num_imgs + num_videos + 
                         global_rate_positive_words + global_subjectivity + 
                         is_weekend, 
               data = train, 
               method = "gbm", 
               trControl = trainControl(method = "cv", number = 5), 
               preProcess = c("center", "scale"),
               tuneGrid = data.frame(expand.grid(n.trees = c(25,50,100,150,200), 
                                                 interaction.depth = 1:4,
                                                 shrinkage = 0.1,
                                                 n.minobsinnode = 10)),
               verbose = FALSE)
boostedTfit
boostedpre <- predict(boostedTfit, newdata = test) 
boosted <- postResample(boostedpre, test$shares)
boosted
```

```{r comparsion of models, echo=TRUE, eval=TRUE}
allRMSE <- tibble(lm1[1], lm2[1], rf[1], boosted[1])
names(allRMSE) <- c("LinearRegression1", "LinearRegression2", "RandomForest", "BoostedTree")
RMSElong <- allRMSE %>%
  pivot_longer(cols = 1:4, names_to = "Model", values_to = "RMSE")
RMSE_sort <- RMSElong %>% 
  arrange(RMSE)
RMSE_sort[1,]
```

The result is the best model and its RMSE.  

# Automation  

```{r automation, echo=TRUE,eval=FALSE}
channels <- c("data_channel_is_lifestyle", "data_channel_is_entertainment", "data_channel_is_bus", "data_channel_is_socmed", "data_channel_is_tech", "data_channel_is_world")
# Create file names
name <- c("Lifestyle", "Entertainment", "Business", "SocialMedia",
          "Tech", "World")
output_file <- paste0(name, "Analysis.md")
# Create a list for each channel with just channel name parameter
params <- lapply(channels, FUN = function(x){
  list(Channels = x)
})
# Put into a data frame
reports <- tibble::tibble(output_file, params)
apply(reports, MARGIN = 1, FUN = function(x) {
  rmarkdown::render(input = "ST558_Project2_Group10.Rmd", 
                    output_format = "github_document", 
                    output_file = x[[1]], 
                    params = x[[2]], 
                    output_options = list(html_preview = FALSE)) 
})
```

