---
title: "ST558 Project2"
author: "Shan Luo, Chengxi Zhou"
date: '2022-07-03'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
library(matrixStats)
library(ggplot2)
```

```{r read in and subset data, echo=TRUE, eval=TRUE}
LifestyleNews <- read_csv("OnlineNewsPopularity.csv", show_col_types = FALSE) %>%
  filter(data_channel_is_lifestyle == 1) %>%
  select(shares, n_tokens_content, num_imgs, num_videos,
         global_rate_positive_words, global_subjectivity, is_weekend)
LifestyleNews$is_weekend <- factor(LifestyleNews$is_weekend)
# Create binary shares variable
binshares <- c()
for (i in seq_len(nrow(LifestyleNews))){
  if (LifestyleNews$shares[i] >= 1400){
    binshares[i] <- 1
  }else{
    binshares[i] <- 0
  }
}
LifestyleNews$binshares <- factor(binshares)
```

```{r check correlation, echo=TRUE, eval=TRUE}
cor(select(LifestyleNews ,shares, n_tokens_content, num_imgs, num_videos,
           global_rate_positive_words, global_subjectivity))
```

All of numeric variables have low correlations.  

```{r split train and test data, echo=TRUE, eval=TRUE}
set.seed(1)
trainIndex <- createDataPartition(LifestyleNews$shares, p = 0.7, list = FALSE)
train <- LifestyleNews[trainIndex, ]
test <- LifestyleNews[-trainIndex, ]
# Train and test set with continuous shares
lmtrain <- train %>% select(-binshares)
lmtest <- test %>% select(-binshares)
# Train and test set with binary shares
bintrain <- train %>% select(-shares)
bintest <- test %>% select(-shares)
```

# EDA  

## Numeric Summary Table  

```{r numeric summaries, echo=TRUE, eval=TRUE}
apply(X = select(train, shares:global_subjectivity), MARGIN = 2,
      FUN = function(x) {
        summaries <- c(min(x), mean(x), median(x), max(x), sd(x), IQR(x))
        names(summaries) <- c("Min", "Mean", "Median", "Max", "Sd", "IQR")
        summaries
      })
```

From numeric summary table, the response **shares** may have right-skewed distribution. The standard deviation of **shares** is unusually high, indicating that there are outliers. The n_tokens_content, num_imgs, and num_videos predictors may have right-skewed distributions. The rate_positive_words and global_subjectivity may have the symmtric distribution.  

## Contingency Table  

```{r contingency table, echo=TRUE, eval=TRUE}
table(train$is_weekend)
table(train$binshares)
table(train$is_weekend, train$binshares)
```

From the contingency table, there are much more articles published on weekday.  

## Bar Plot  

```{r barplot, echo=TRUE, eval=TRUE}
g <- ggplot(data = train, aes(x = is_weekend))
g + geom_bar(fill = "Red", color = "Blue") +
  labs(title = "Bar Plot of is_weekend")

g <- ggplot(data = train, aes(x = binshares))
g + geom_bar(fill = "Red", color = "Blue") +
  labs(title = "Bar Plot of Binshares")
```

```{r boxplot, echo=TRUE, eval=TRUE}
g <- ggplot(data = train, aes(x = is_weekend, y = shares))
g + geom_boxplot() +
  geom_jitter(aes(color = is_weekend)) + 
  labs()
```


```{r histograms, echo=TRUE, eval=TRUE}
g <- ggplot(data = train, aes(x = shares))
g + geom_histogram(bins = 30, aes(fill = is_weekend)) +
  labs(x = "Number of Shares",
       title = "Histogram of Shares") +
  scale_fill_discrete(name = "Weekend Published", labels = c("No", "Yes"))
```

```{r scatterplot, echo=TRUE, eval=TRUE}
g <- ggplot(data = train, aes(x = n_tokens_content, y = shares))
g + geom_point(aes(color = is_weekend)) +
  geom_smooth(method = "lm")
```

# Modeling  

```{r linear regression, echo=TRUE, eval=TRUE}
lmfit <- train(shares ~.,
               data = lmtrain,
               method = "lm")
lmfit
lmpred <- predict(lmfit, newdata = lmtest) 
postResample(lmpred, lmtest$shares)
```

```{r random forest, echo=TRUE, eval=TRUE}
RFfit <- train(shares ~ ., 
                data = lmtrain, 
                method = "rf", 
                trControl = trainControl(method = "cv", number = 5), 
                preProcess = c("center", "scale"),
               tuneGrid = data.frame(mtry = 1:6))
RFfit
RFpred <- predict(RFfit, newdata = lmtest) 
postResample(RFpred, lmtest$shares)
```

```{r}

```

